### 2020-03-17 thesis_template/2_RelatedWork/relatedWork.tex
added documentation
documented

\section{Memory Management in Rust}
\label{sec:history}
\input{2_RelatedWork/model/rust_memory_management}

\section{State of the Art}
\label{sec:history}
\input{2_RelatedWork/model/spark}

@@

\section{Experiment of Memory Allocation}
\label{sec:history}
This experiment is to test how static and dynamic memory allocation of Java and Rust behave. For the assessment, Element addition to ArrayList in the case of Java and Vector in the case of Rust 
is emploied here. These data stractures are resizable and we have control to set initial size. There are two parameters; initial size of ArrayList or Vector and thier final size after additions of elements. 
We are interested in the impact to memory allocation by initial allocation and expantion to runtime.

First, an ArrayList and an Vector are created with specified initial size. Then, in a loop, a element is added for each iteration until the size of the ArrayList or Vector get the specified final size. 
Each data structure has a different resizing strategy. When an ArrayList hits current limit of its size and expands the limit, it doubles the current size.  
While Vector does not have specific strategy for its resizing, the expantions of size of both ArrayList and Vector might affect the digradation of runtime performance.

To perform benchmarks, we use Java Microbenchamrk Hardness (JMH) for Java and Criterion for Rust. The benchmark time is calculated mean from several iterarions. 
We warm up before the execution. The parameters are set in combination of initial size, 10, 100, 1000, and 10000 and final size, 9, 99, 999, 9999. 
The results are shown in Figure 2-4 and 2-5.

Discussions here are separated to two cases: when initial size is set bigger than final size and when final size is set bigger than initial size. 
For ArrayList in Java, it shows performance significant degradation when the initial size is bigger than final size. When we create ArrayList with intial size of 1000 and add 99 elements 
to the ArrayList, the the average execution time is 1125 ns. However, the average excution time when we create ArrayList with intial size of 100 and add 99 elements is 623 ns. 
This degration is caused by the cost of initializing the large array. On the other hand, Rust Vector does not have significant cost for the initialization of size compared to the cost of addition of elements. 
In the case of Vector with initial size of 1000 and add 99 elements to the Vector, the average excution is 320 ns. In the case of the initial size of 100 and 99 elements additions, 
the average excution is 279 ns. This is such small degradation compared to element addition in Vector.

In ArrayList when the initial size is smaller than the final size, there can be degradation in performance. When its initial size and final size are set 100 and 999 respectively, 
the average execution time is 10884 ns. However when its initial size and final size are set 1000 and 999 respectively, the average execution time is 4250 ns. 
This result shows the degradation of in performance caused by the copy of existing elements into newly allocated array whose size is double of the last array.

This characteristic can be seen in the case of Vector in Rust. Vector with initial size 100 and final size 999 performs the average excution time 3514 ns, 
but one with initial size 1000 and final size 999 performs 2723 ns. When the Vector reachs its capacity, it allocates a larger buffer and copies the present elements to it.
This cost results in the degradation of the average excution time.

\begin{figure}[htb]
    \includegraphics[width=15cm]{java_arraylist.eps}
    \caption{Memory allocation of ArrayList in Java}
    \label{fig:Sampling}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_vector.eps}
    \caption{Memory allocation of Vector in Rust}
    \label{fig:Sampling}
\end{figure}

\section{Experiment of Memory Allocation for Different Element Type.}
\label{sec:history}
This experiment is to test how static and dynamic memory allocation of Java and Rust behave. For the assessment, Element addition to ArrayList in the case of Java and Vector in the case of Rust 
is emploied here. These data stractures are resizable and we have control to set initial size. There are two parameters; initial size of ArrayList or Vector and thier final size after additions of elements. 
We are interested in the impact to runtime performance by initializing memory allocation and dynamicaly allocating memory space.

First, an ArrayList and an Vector are created with specified initial size. Then, in a loop, a element is added for each iteration until the size of the ArrayList or Vector get the specified final size. 
Each data structure has a different resizing strategy. When an ArrayList hits current limit of its size and expands the limit, it doubles the current size.  
While Vector does not have specific strategy for its resizing, the expantions of size of both ArrayList and Vector might affect the digradation of runtime performance.

Second, four types of element are used for elements addition to each data structure: integer, array of charactors, string, and Customer object. 
Assamption is that there would be different behavior between element additions of dynamically resizable and static size objects. 
Customer object has three fields. These fields are total order, weight of order, and zip code whose types are integer (i32 in rust), double (f32 in rust), and string respectively.
Figure 2-6 and 2-7 are representations of customer objects in Java and Rust.

Figure 2-10, 2-11, 2-12, and 2-13 represent the result of the experiments. For both data structures, integer elements addition shows the fastest runtime among all object types. 
This is because the compilers know each integer need 4 bytes to be stored in memory so that the space for memory that should be allocated is easily inspected. 
For the same reason, the initialization of data structures whose elements are integers alaways improves runtim performance.

The elements addition of strings and array of character behave similally among each languages. These two types of elements addition perform the similar speed and significantly slower than integer addition. 
Customer object addition is the slowest in Java. However, in Rust the addition of Customer object is the second fastest among all element typs.
The impacts of initialization of Java ArrayList vary among elemet types. On the other hand, the initailization Rust Vector always improves runtime performance for any of 4 types of elements addition.

\begin{figure}[htb]
    \begin{lstlisting}
        class Customer {
            int totalOrder;
            double weightOrder;
            String zipCode;
        }
    \end{lstlisting}
    \caption{Representation of Customer object in Java.}
    \label{fig:Sampling}    
\end{figure}

\begin{figure}[htb]
    \begin{lstlisting}
        struct Customer {
            total_order: i32,
            weight_order: f32,
            zip_code: String,
        }
    \end{lstlisting}
    \caption{Representation of Customer object in Rust.}
    \label{fig:Sampling}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{java_arraylist_log.eps}
    \caption{Memory allocation of Java ArrayList}
    \label{fig:Sampling}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_vector_log.eps}
    \caption{Memory allocation of Rust Vector}
    \label{fig:Sampling}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{java_arraylist_difference.eps}
    \caption{Difference of Memory allocation of Java ArrayList between Non-initialization and Initialization}
    \label{fig:Sampling}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_arraylist_difference.eps}
    \caption{Difference of Memory allocation of Rust Vector between Non-initialization and Initialization}
    \label{fig:Sampling}
\end{figure}

@@

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_merge_sort.eps}
    \caption{mergesort}
    \label{fig:Sampling}
\end{figure}

>>>>

\label{sec:history}
\input{2_RelatedWork/model/spark}

@@

@@

### 2020-03-17 thesis_template/3_Body/model/merge_sort_experiment.tex
documented

This experiment is to assess importance of careful memory management in multithreading of Rust. 
We implement merge-sort algorithm in three different ways. One is sharing source vector with Arc. 
Another is passing slice of source vector to child thread. The other is LinkedList implementation other than vector.

@@

For sharing data implementaion, channel with sending data is used for multithreading method, because we want to ensure children threads return values before parent thread proceed execution. 
For passing slice implementation, we use scope method to enable children threads to receive reference from their parent ensuring the same purpose of sending data. 
The linkedlist implementation is different from another two. This algorithm is inplace sorting so that it does not de/allocate memory during execution.

The comparison among sharing data and passing slice implementations is to see the impact of Arc, Atomic reference conuting, to runtime performance.
These two implementations are doing the same operations other than using or not using Arc to share data. 
This comparison can effectively show how important careful memory management is in multithreading computation in Rust.

One concern of these vector implementation is that type of elements of vector. To create merged vector in every merging phase, we need to get element value of vector at base case.
To acquire the value of element, we need reference counting as element or copying the value of object. Both does not seem perform the best in other languages.

The comparison among vector and linkedlist implementations can show trade-off between contiguous memory access and inplace non memory de/allocation. 
Experiment with Java linkedlist implementation can be interesting, because Java GC is severe problem when number of object is large.

>>>>

This experiment is to assess importance of careful memory management in multithreading of Rust. 
We implement merge-sort algorithm in two different ways. One is sharing source vector with Arc. 
The other is passing slice of source vector to child thread.

@@

For sharing data implementaion, channel with sending data is used for multithreading method, because we want to ensure children threads return values before parent thread proceed execution. 
For passing slice implementation, we use scope method to enable children threads to receive reference from their parent ensuring the same purpose of sending data.

Experiment performed is the comparison among sharing data and passing slice implementations to see the impact of Arc, Atomic reference conuting, to runtime performance.
These two implementations are theoritically the same operations other than using or not using Arc to share data. 
This comparison can effectively show how important careful memory management is in multithreading computation in Rust. 
In another word, how atomic reference counting can cost for computation in Rust programming.

The figure shows the result of 
An atomic reference count of the shareable vector is passed to every recursion steps and

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_merge_sort.eps}
    \caption{Runtime of Sortting Elements of Customer Vector}
    \label{fig:Sampling}
\end{figure}

The other is LinkedList implementation other than vector.
The linkedlist implementation is different from another two. This algorithm is inplace sorting so that it does not de/allocate memory during execution. 
The comparison among vector and linkedlist implementations can show trade-off between contiguous memory access and inplace non memory de/allocation. 
Experiment with Java linkedlist implementation can be interesting, because Java GC is severe problem when number of object is large.
